---
title: "Sales Report"
author: "Buy & Sell Goods App"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: xelatex
    keep_tex: false
    pandoc_args: ["--variable", "geometry:landscape"]
params:
  sales_data: NULL
  storage_data: NULL
  cogs_summary: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(lubridate)
library(knitr)
library(kableExtra)
library(writexl)
library(readxl)
library(forecast)
```

**Report Generated:** `r params$report_time`

## Executive Summary

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(params$sales_data) && !is.null(params$storage_data) &&
    nrow(params$sales_data) > 0 && nrow(params$storage_data) > 0) {
  total_sales <- sum(params$sales_data$Total, na.rm = TRUE)
  total_tax <- total_sales * 0.10
  net_revenue <- total_sales - total_tax

  relevant_items <- unique(params$sales_data$Item)
  total_items_sold <- length(relevant_items)
  total_units_sold <- sum(params$sales_data$Quantity, na.rm = TRUE)

  knitr::asis_output(paste0(
    "**Executive Summary**\n\n",
    "- **Total Gross Sales**: Rp", format(round(total_sales, 2), big.mark = ",", scientific = FALSE), "\n",
    "- **Total PPN Collected (10%)**: Rp", format(round(total_tax, 2), big.mark = ",", scientific = FALSE), "\n",
    "- **Total Net Revenue**: Rp", format(round(net_revenue, 2), big.mark = ",", scientific = FALSE), "\n",
    "- **Unique Items Sold**: ", total_items_sold, "\n",
    "- **Total Units Sold**: ", total_units_sold
  ))
} else {
  cat("Not enough data to generate executive summary.")
}
```

## Total Sales Summary

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(params$sales_data) && nrow(params$sales_data) > 0) {
  total_sales <- sum(params$sales_data$Total, na.rm = TRUE)
  total_tax <- total_sales * 0.10
  net_revenue <- total_sales - total_tax

  summary_df <- data.frame(
    "Total Gross Sales" = sprintf("%.2f", total_sales),
    "Total PPN Collected (10%)" = sprintf("%.2f", total_tax),
    "Total Net Revenue" = sprintf("%.2f", net_revenue)
  )
  kable(summary_df, caption = "Total Sales Revenue Summary")
} else {
  cat("No sales data available.")
}
```

## Monthly Sales Overview

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(params$sales_data) && nrow(params$sales_data) > 0) {
  monthly_summary <- params$sales_data %>%
    mutate(Month = format(as.POSIXct(Date), "%Y-%m")) %>%
    group_by(Month) %>%
    summarise(
      `Total Sales` = sum(Total, na.rm = TRUE),
      `Units Sold` = sum(Quantity, na.rm = TRUE),
      .groups = "drop"
    )

  kable(monthly_summary, caption = "Monthly Sales Summary", booktabs = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Summary of Sold Items by Date and Time

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(params$sales_data) && nrow(params$sales_data) > 0) {
  summary_data <- params$sales_data %>%
    mutate(Date = format(as.POSIXct(Date), "%Y-%m-%d %H:%M:%S")) %>%
    group_by(Date, Item) %>%
    summarise(
      `Quantity Sold` = sum(Quantity, na.rm = TRUE),
      `Gross Revenue` = sum(Total, na.rm = TRUE),
      `Tax Collected` = round(`Gross Revenue` * 0.10, 2),
      `Net Revenue` = round(`Gross Revenue` * 0.90, 2),
      `Price per Unit` = round(mean(Price, na.rm = TRUE), 2),
      .groups = "drop"
    ) %>%
    arrange(Date)

  kable(summary_data, caption = "Summary of Sold Items by Date")
}
```

## Sales by Hour of Day

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(params$sales_data) && nrow(params$sales_data) > 0) {
  hourly_summary <- params$sales_data %>%
    mutate(Hour = format(as.POSIXct(Date), "%H")) %>%
    group_by(Hour) %>%
    summarise(`Total Sales` = sum(Total, na.rm = TRUE), .groups = "drop") %>%
    mutate(Hour = as.numeric(Hour)) %>%
    arrange(Hour)

  ggplot(hourly_summary, aes(x = Hour, y = `Total Sales`)) +
    geom_col(fill = "#0072B2") +
    labs(title = "Sales by Hour of Day", x = "Hour", y = "Total Sales (Rp)") +
    theme_minimal()
}
```

## Detailed Sales Log

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(params$sales_data) && nrow(params$sales_data) > 0) {
  detailed_log <- params$sales_data %>%
    mutate(
      Timestamp = format(as.POSIXct(Date), "%Y-%m-%d %H:%M:%S"),
      Tax = round(Total * 0.10, 2),
      `Net Revenue` = round(Total * 0.90, 2)
    ) %>%
    select(Timestamp, Item, Quantity, Price, Total, Tax, `Net Revenue`)
  
  kable(detailed_log, caption = "Detailed Sales Log")
}
```

\newpage

## Storage Records and Purchases vs Sales (with Net Profit Calculation)

```{r full-summary, echo=FALSE, message=FALSE, warning=FALSE}
sales_data <- params$sales_data
storage_data <- params$storage_data

if (is.null(sales_data) || is.null(storage_data) || nrow(sales_data) == 0 || nrow(storage_data) == 0) {
  cat("## Not enough data to generate full inventory summary.")
} else {
  combined_summary <- list()

  for (i in 1:nrow(storage_data)) {
    row <- storage_data[i, ]
    combined_summary[[length(combined_summary) + 1]] <- data.frame(
      Date = format(as.POSIXct(row$Date), "%Y-%m-%d %H:%M:%S"),
      Item = row$Item,
      Type = "Stock Added",
      Quantity = row$Stock_Added,
      Price = row$Price,
      Revenue = NA_real_,
      Cost_per_Unit = round(row$Price, 2),
      COGS = NA_real_,
      Tax = NA_real_,
      Net_Revenue = NA_real_,
      Net_Profit = NA_real_,
      stringsAsFactors = FALSE
    )
  }

  for (i in 1:nrow(sales_data)) {
    sale_row <- sales_data[i, ]
    sale_date <- as.POSIXct(sale_row$Date)
    sale_qty <- sale_row$Quantity
    gross_total <- sale_row$Total
    tax <- 0.10 * gross_total
    net_revenue <- gross_total - tax

    relevant_stock <- storage_data %>%
      filter(Item == sale_row$Item & as.POSIXct(Date) <= sale_date)

    if (nrow(relevant_stock) > 0) {
      total_qty <- sum(relevant_stock$Stock_Added, na.rm = TRUE)
      total_cost <- sum(relevant_stock$Stock_Added * relevant_stock$Price, na.rm = TRUE)
      avg_cost <- ifelse(total_qty > 0, total_cost / total_qty, 0)
    } else {
      avg_cost <- 0
    }

    cogs <- sale_qty * avg_cost
    net_profit <- net_revenue - cogs

    combined_summary[[length(combined_summary) + 1]] <- data.frame(
      Date = format(sale_date, "%Y-%m-%d %H:%M:%S"),
      Item = sale_row$Item,
      Type = "Sale",
      Quantity = sale_qty,
      Price = sale_row$Price,
      Revenue = round(gross_total, 2),
      Cost_per_Unit = round(avg_cost, 2),
      COGS = round(cogs, 2),
      Tax = round(tax, 2),
      Net_Revenue = round(net_revenue, 2),
      Net_Profit = round(net_profit, 2),
      stringsAsFactors = FALSE
    )
  }

  final_df <- bind_rows(combined_summary) %>%
    arrange(as.POSIXct(Date))

  final_df <- final_df %>%
    rename_with(~ gsub("_", " ", .x))

  stock_added_table <- final_df %>%
    filter(Type == "Stock Added") %>%
    select(Date, Item, Quantity, `Cost per Unit`)

  sales_table <- final_df %>%
    filter(Type == "Sale") %>%
    select(Date, Item, Quantity, Price, Revenue, Tax, `Net Revenue`, `Cost per Unit`, COGS, `Net Profit`)
}
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("stock_added_table")) {
  kable(stock_added_table, caption = "Report: Stock Added", booktabs = TRUE, longtable = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("sales_table")) {
  kable(sales_table, caption = "Report: Sales Summary", booktabs = TRUE, longtable = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Combined Inventory Summary

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)
library(scales)

if (exists("final_df") && nrow(final_df) > 0) {
  final_df_clean <- final_df %>%
    mutate(across(where(is.numeric), ~ round(., 2)))

  final_df_clean[is.na(final_df_clean)] <- ""

  display_df <- final_df_clean %>%
    mutate(across(
      .cols = where(~ all(grepl("^[0-9.,]*$", .x) | .x == "")),
      .fns = ~ ifelse(.x == "", "", format(as.numeric(.x), big.mark = ",", scientific = FALSE))
    ))

  kable(display_df, caption = "Combined Inventory Summary", booktabs = TRUE, longtable = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Estimated Ending Inventory

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("final_df") && nrow(final_df) > 0) {
  stock_levels <- final_df %>%
    group_by(Item) %>%
    summarise(
      `Stock Added` = sum(ifelse(Type == "Stock Added", Quantity, 0), na.rm = TRUE),
      `Stock Sold` = sum(ifelse(Type == "Sale", Quantity, 0), na.rm = TRUE),
      `Stock Remaining` = `Stock Added` - `Stock Sold`,
      .groups = "drop"
    ) %>%
    arrange(desc(`Stock Remaining`))

  kable(stock_levels, caption = "Estimated Ending Inventory", booktabs = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Items Sold Without Stock Record

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(params$sales_data) && !is.null(params$storage_data)) {
  missing_stock <- setdiff(unique(params$sales_data$Item), unique(params$storage_data$Item))

  if (length(missing_stock) > 0) {
    knitr::asis_output(paste("**Warning:** The following items were sold but never recorded as stocked:\n\n- ", paste(missing_stock, collapse = "\n- ")))
  }
}
```

## Low Stock Alert (Stock < 5)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Set the threshold for low stock
low_stock_threshold <- 3

if (exists("stock_levels") && nrow(stock_levels) > 0) {
  low_stock <- stock_levels %>%
    filter(`Stock Remaining` <= low_stock_threshold & `Stock Remaining` > 0) %>%
    arrange(`Stock Remaining`)

  if (nrow(low_stock) > 0) {
    kable(low_stock, caption = paste("Items with Stock ≤", low_stock_threshold), booktabs = TRUE) %>%
      kable_styling(latex_options = c("striped", "scale_down"))
  } else {
    knitr::asis_output("**No items are below the low stock threshold.**")
  }
} else {
  knitr::asis_output("**No stock data available for low stock alert.**")
}
```

## Sales Over Time

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- params$sales_data
if (nrow(df) > 0) {
  df$Date <- as.Date(as.POSIXct(df$Date))
  
  summary_df <- df %>%
    group_by(Date) %>%
    summarise(Total_Sales = sum(Total, na.rm = TRUE), .groups = "drop")
  
  ggplot(summary_df, aes(x = Date, y = Total_Sales)) +
    geom_line(color = "#0072B2", size = 0.5) +
    geom_point(color = "#0072B2", size = 2) +
    geom_smooth(method = "loess", se = FALSE, color = "darkorange", linetype = "dashed", size = 1) +
    labs(
      title = "Total Sales Over Time",
      x = "Date",
      y = "Total Sales (Rp)"
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma)
}
```

## Sales Forecast (Next 7 Days)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(forecast)
library(ggplot2)
library(scales)

min_data_points <- 14

df <- params$sales_data
if (nrow(df) > 0) {
  df$Date <- as.Date(as.POSIXct(df$Date))

  summary_df <- df %>%
    group_by(Date) %>%
    summarise(Total_Sales = sum(Total, na.rm = TRUE), .groups = "drop") %>%
    arrange(Date)

  if (nrow(summary_df) >= min_data_points) {
    ts_data <- ts(summary_df$Total_Sales, frequency = 7)  # assume weekly seasonality

    model <- auto.arima(ts_data)
    forecasted <- forecast(model, h = 7)  # forecast next 7 days

    autoplot(forecasted) +
      labs(title = "7-Day Sales Forecast",
           x = "Date",
           y = "Forecasted Sales (Rp)") +
      theme_minimal() +
      scale_y_continuous(labels = comma)
  } else {
    knitr::asis_output(paste0(
      "**Not enough data to generate forecast.**\n\n",
      "- Minimum required: ", min_data_points, " days of sales\n",
      "- Available: ", nrow(summary_df), " days"
    ))
  }
} else {
  knitr::asis_output("**Sales data not available for forecasting.**")
}
```

## Profit Margin Over Time

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- params$cogs_summary
if (!is.null(df) && nrow(df) > 0) {
  sales_df <- df[df$Type == "Sale", ]
  
  if (nrow(sales_df) > 0) {
    sales_df$Date <- as.Date(sales_df$Date)
    sales_df$Profit_Margin <- with(sales_df, ifelse(Revenue > 0, `Net Profit` / Revenue, NA))
    
    summary_df <- sales_df %>%
      group_by(Date) %>%
      summarise(Avg_Margin = mean(Profit_Margin, na.rm = TRUE), .groups = "drop")
    
    ggplot(summary_df, aes(x = Date, y = Avg_Margin)) +
      geom_line(color = "forestgreen", size = 0.5) +
      geom_point(color = "forestgreen", size = 2) +
      geom_smooth(method = "loess", se = FALSE, color = "orange", linetype = "dashed") +
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
      labs(
        title = "Average Profit Margin Over Time",
        x = "Date",
        y = "Profit Margin"
      ) +
      theme_minimal()
  }
}
```

## Cumulative Sales and Profit Over Time

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(params$sales_data) > 0) {
  cum_data <- final_df %>%
    filter(Type == "Sale") %>%
    mutate(Date = as.Date(Date)) %>%
    arrange(Date) %>%
    group_by(Date) %>%
    summarise(
      Revenue = sum(Revenue, na.rm = TRUE),
      Profit = sum(`Net Profit`, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      `Cumulative Revenue` = cumsum(Revenue),
      `Cumulative Profit` = cumsum(Profit)
    )

  ggplot(cum_data, aes(x = Date)) +
    geom_line(aes(y = `Cumulative Revenue`), color = "blue", size = 0.5) +
    geom_line(aes(y = `Cumulative Profit`), color = "darkgreen", size = 0.5, linetype = "dashed") +
    labs(
      title = "Cumulative Revenue and Profit Over Time",
      x = "Date", y = "Value (Rp)"
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma)
}
```

## Summary Based on Per Item

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("final_df") && nrow(final_df) > 0) {
  final_numeric <- final_df %>%
    filter(Type == "Sale") %>%
    mutate(across(
      .cols = intersect(c("Quantity", "Revenue", "Tax", "Net Revenue", "COGS", "Net Profit"), names(.)),
      .fns = ~ as.numeric(.)
    ))

  per_item_summary <- final_numeric %>%
    group_by(Item) %>%
    summarise(
      `Total Quantity Sold` = sum(Quantity, na.rm = TRUE),
      `Gross Revenue` = sum(Revenue, na.rm = TRUE),
      `PPN Collected` = sum(Tax, na.rm = TRUE),
      `Net Revenue` = sum(`Net Revenue`, na.rm = TRUE),
      `Total COGS` = sum(COGS, na.rm = TRUE),
      `Total Net Profit` = sum(`Net Profit`, na.rm = TRUE),
      .groups = "drop"
    )

  total_row <- per_item_summary %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(Item = "Total")

  per_item_summary <- bind_rows(per_item_summary, total_row)

  kable(per_item_summary, caption = "Sales Summary by Item", booktabs = TRUE, longtable = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Sales by Category

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if ("Category" %in% colnames(params$sales_data)) {
  category_summary <- params$sales_data %>%
    group_by(Category) %>%
    summarise(
      `Total Sales` = sum(Total, na.rm = TRUE),
      `Units Sold` = sum(Quantity, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(desc(`Total Sales`))

  kable(category_summary, caption = "Sales Summary by Category", booktabs = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Per Unit Sales & Cost Analysis

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("final_df")) {
  unit_economics <- final_df %>%
    filter(Type == "Sale") %>%
    group_by(Item) %>%
    summarise(
      `Avg Sale Price` = round(mean(Price, na.rm = TRUE), 2),
      `Avg Cost per Unit` = round(mean(`Cost per Unit`, na.rm = TRUE), 2),
      `Avg Net Profit per Unit` = round(mean(`Net Profit` / Quantity, na.rm = TRUE), 2),
      .groups = "drop"
    )

  kable(unit_economics, caption = "Unit Economics per Item", booktabs = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Sales Priced Below Cost

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("final_df")) {
  below_cost_sales <- final_df %>%
    filter(Type == "Sale", Price < `Cost per Unit`) %>%
    arrange(desc(`Cost per Unit` - Price))

  if (nrow(below_cost_sales) > 0) {
    kable(below_cost_sales, caption = "Sales Priced Below Cost", booktabs = TRUE) %>%
      kable_styling(latex_options = c("striped", "scale_down"))
  }
}
```

## Top 5 Best selling items

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("per_item_summary")) {
  top_items <- per_item_summary %>%
    filter(Item != "Total") %>%
    arrange(desc(`Total Net Profit`)) %>%
    slice_head(n = 5)

  kable(top_items, caption = "Top 5 Most Profitable Items", booktabs = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```

## Bottom 5 Least Profitable Items

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if (exists("per_item_summary")) {
  bottom_items <- per_item_summary %>%
    filter(Item != "Total") %>%
    arrange(`Total Net Profit`) %>%
    slice_head(n = 5)

  kable(bottom_items, caption = "Bottom 5 Least Profitable Items", booktabs = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"))
}
```
